---
title: "R Interface to CloudML"
output: 
  rmarkdown::html_vignette: 
    readme: true
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```

[![Build Status](https://travis-ci.org/rstudio/cloudml.svg?branch=master)](https://travis-ci.org/rstudio/cloudml) [![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/cloudml)](https://cran.r-project.org/package=cloudml)

## Overview

The **cloudml** package provides an R interface to [Google Cloud Machine Learning](https://cloud.google.com/ml-engine/). CloudML makes it easy to run training applications built with R packages like [tensorflow](https://tensorflow.rstudio.com/) and [keras](https://keras.rstudio.com/) in the cloud.

CloudML provides a cost-effective way to use GPUs for training models, since you only pay for the GPU compute time that your training jobs actually consume (as opposed to investing in a dedicated GPU machine either locally or in the cloud). CloudML also provides access to powerful scalable computing facilities when you need them.

## Prerequisites

#### Google Cloud Account

Before you can begin training models with CloudML you need to have a **Google Cloud Account**. If you don't already have an account you can create one at <https://console.cloud.google.com>.

#### CloudML R Package

Next, install the cloudml R package from GitHub as follows:

```r
devtools::install_github("rstudio/cloudml")
```

#### Google Cloud SDK

Finally, you need to install the **Google Cloud SDK**, a set of utilties that enable you to interact with your Google Cloud account from within R. You can install the SDK using the `gcloud_install()` function:

```{r}
gcloud_install()
```

Once you've completed these steps you are ready to train models with CloudML!

## Training on CloudML

To train models on CloudML, first work on them locally (perhaps with a smaller sample of your dataset) until your training script is working correctly. Then, call the `cloudml_train()` function, specifying the R script to execute for training:

```{r}
job <- cloudml_train("train.R")
```

Note that all of the files within the current working directory will be bundled up and sent along with the script.

When using RStudio, a terminal window is used to stream the logs and download the job when it finializes.

When not using RStudio, you can ask the R session to wait for training to complete, and pull the generated models back to your local filesystem with:

```{r eval=FALSE}
job_collect(job)
```

The R session will wait and continue polling Google Cloud until your application has finished running; if the application trained successfully, then the trained models will be copied to disk under the `runs` directory.

Once the job is collected, you can use [tfruns](https://tensorflow.rstudio.com/tools/tfruns/) to view the results of the last run:

```{r eval=FALSE}
library(tfruns)
view_run()
```

