---
title: "R Interface to CloudML"
output: 
  rmarkdown::html_vignette: 
    readme: true
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```

[![Build Status](https://travis-ci.org/rstudio/cloudml.svg?branch=master)](https://travis-ci.org/rstudio/cloudml) [![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/cloudml)](https://cran.r-project.org/package=cloudml)

## Overview

The **cloudml** package provides an R interface to [Google Cloud Machine Learning](https://cloud.google.com/ml-engine/). CloudML makes it easy to run training applications built with R packages like [tensorflow](https://tensorflow.rstudio.com/) and [keras](https://keras.rstudio.com/) in the cloud. 

Cloud ML Engine manages the computing resources that your training job needs to run, so you can focus more on your model than on hardware configuration or resource management. Your use of the training service is bounded by pricing and resource quota policy, but within those bounds, you have access to powerful scalable computing.

## Prerequisites

### Google Cloud Account

Before you can begin training models with CloudML you need to have a **Google Cloud Account**. If you don't already have an account you can create one at <https://console.cloud.google.com>.

### CloudML R Package

Next, install the cloudml R package from GitHub as follows:

```r
devtools::install_github("rstudio/cloudml")
```

### Google Cloud SDK

Finally, you need to install the **Google Cloud SDK**, a set of utilties that enable you to interact with your Google Cloud account from within R. 

You can install the SDK using the `gcloud_install()` function:

```{r}
gcloud_install()
```

Once you've completed these steps you are ready to train models with CloudML!

## Training on CloudML

To train models on CloudML, first work on them locally (perhaps with a smaller sample of your dataset) until your training script is working correctly. Then, call the `cloudml_train()` function, specifying the R script to execute for training:

```{r}
job <- cloudml_train("train.R")
```

Note that all of the files within the current working directory will be bundled up and sent along with the script.

When using RStudio, a terminal window is used to stream the logs and download the job when it finializes.

When not using RStudio, you can ask the R session to wait for training to complete, and pull the generated models back to your local filesystem with:

```{r eval=FALSE}
job_collect(job)
```

The R session will wait and continue polling Google Cloud until your application has finished running; if the application trained successfully, then the trained models will be copied to disk under the `runs` directory.

Once the job is collected, you can use [tfruns](https://tensorflow.rstudio.com/tools/tfruns/) to view the results of the last run:

```{r eval=FALSE}
library(tfruns)
view_run()
```

